Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-02-28 SIR_manuscript.tex
all plots to pdf, flowchart modified

\begin{figure}[!h] 
\begin{center} 
\includegraphics[scale=0.3]{./pix/sir_comp.png}
\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}. Here, the disease-based status of a compartment $X$, where $X \in \{S,I,R\}$, is combined with the testing-based status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively. Also,  $\Lambda$ is the force of infection with definition in Eq.~\eqref{Lambda}, $\gamma$ is the recovery rate, $\omega$ is the rate of test return, $F_X$ and $p_X$ represent the \percap testing rate and the probability of positive tests, respectively, for compartment $X$. For further description of the parameters see \cref{tab:params}. \ali{(i) will change the subs to roman font, (ii) not sure if ref color is helpful in u compartment?}
\label{fig:flowchart}}
\end{center} 
\end{figure}

Table~\ref{tab:params} defines the model parameters, which are generally straightforward
\percap flows between compartments, or modifiers to these flow rates. The \percap testing intensity is defined as the number of daily tests taken in a population of size $N_0$. 
The novel component of the model comes in through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively. For example, $W_I/W_S=2$ means that infected individuals are tested at twice the \percap rate of susceptible individuals.

In order to link to more applied models, i.e., including both epidemiological and testing processes, we constructed this model so we could specify the total \percap testing rate. We do this by defining the weighted size of the testing pool $W = W_S S\_u + W_I I\_u + W_R R\_u$, and calculating a scaling parameter for testing as:
\begin{equation}
\label{sigma}
\sigma = \frac{\rho N_0}{W}.
\end{equation}
Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma W_X$, where $X \in \{S,I,R\}$. 
For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.  Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can (if large enough) exhaust the pool of people available for testing, leading to a singularity when no one is left untested.  Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (we discuss an adjustment to the model that solves this problem in the appendix).

@@

\begin{table}[htp]
\centering
\begin{tabular}{|c|L{2in}cc|} \hline
  Symbol & Description & Unit & Value \\ \hline
  $N_0$     & Total population size & people & $10^6$ \\ \hline
  $\omega$  & Rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
  $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
  $\eta\_w$  & Relative probability of transmission for ``waiting'' individuals & - & - \\ \hline
  $\eta\_c$  & Relative probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
  $\beta$ & Transmission rate & 1/day & 0.39 \\ \hline
  $\Lambda$ & Force of infection & 1/day & - \\ \hline
  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
  $W_S, W_I, W_R$ & Relative testing weights & - &
  \begin{minipage}[t]{0.21\columnwidth}%
    Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
  \end{minipage} \\
  \hline
  \end{tabular}
\caption{\label{tab:params} Parameters of the model \eqref{model}.}
\end{table}

@@

The analytical calculation of the next generation matrix and the derivation of $\Rnum$ was carried out in Maple \citep{maple14} by using simple linear Algebra package. 
  We used R \citep{r} for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ \eqref{R0}. The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing. The parameter values or ranges are presented in Table~\ref{tab:params}. This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal. For simulation purposes we chose $W_S=W_I=W_R=1$ for random testing, and $W_S=0.3$ and $W_I=W_R=1$ for non-random testing strategy. Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.

@@

\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_random.pdf}
\caption{}\label{p.a}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_TTI.pdf}
\caption{}\label{p.b}
\end{subfigure}
\caption{
{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at differnt levels of testing and isolation.}
We numerically evaluate $\Rnum$ \eqref{R0} using the following parameters (listed in Table \ref{tab:params}):
$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\eta\_w \in [0,1]$, $\eta\_c \in [0,1]$, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$. Contours of $\Rnum$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $W_S=W_I=W_R=1$ and (b) targeted testing where $W_S=0.3$ and $W_I=W_R=1$. The black solid line in each panel represents the critical contour of $\Rnum=1$. \bmb{we still have some collisions in the y-axis tick labels}}
\label{pan}
\end{figure}

@@

We incorporated the compartment-specific relative testing weights, $W_S$, $W_I$ and $W_R$, to model random testing and \emph{targeted} testing strategies. Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $W_I/W_S=3$ and $W_R=W_I$. Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario. 
This part needs to be developed further in future work. 
Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment. 
This can be implied when we would like to quantify and compare the effect of different levels of test focus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management. For example, when people are tested for "screening", the individuals with potential higher mobility, eg. people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with. With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing processes.

@@

\ali{ not sure if we want these 2 following parageraphs, or shrink it into afew short sentences?}
In addition to the favorable-delay-reporting observation of our model \eqref{model}, the model enables us to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$.
To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf}. To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta\_c \ll \eta\_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $W_i \gg W_s$).

@@

% %%%%%%%
\subsection{Taylor Approximation of $\Rnum$ at $\rho=0$ }
The basic reproduction number, $\Rnum$, close to $\rho=0$ can be approximated linearly in $\rho$ by using Taylor approximation. it follows
\begin{equation}
\label{eq:R0appr}
\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta\_w-1)(\gamma W_S+\omega W_I) + (\eta\_c -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
\end{equation}

Also, to analyze the influence of $\omega$ on $\Rnum$, the approximation \eqref{eq:R0appr} was used. Then it is straight forward to have
\begin{equation}
\label{Rom}
\partial{\Rnum}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
\end{equation}
where $a=(\eta\_w-1)W_I-(\eta\_c-1)P_I W_I = ((s-p_I)\eta\_c + (p_I-1)) W_I$, $b=2(\eta\_w-1)\gamma W_S$ and $c=(\eta\_w-1)\gamma^2 W_S$.
Given that $0 \leq \eta\_c\leq \eta\_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.

@@

% %%%%%%%
\subsection{rate of returning tests} 
\ali{needs editig}
The linearization of $\Rnum$ around $\rho=0$ is
\begin{equation}\label{linearization}
\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta\_w-1)(\gamma W_s+\omega W_i) + (\eta\_c -1)P_iW_i \omega^2 \Big). 
\end{equation}
So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
where $a=(\eta\_w-1)W_i-(\eta\_c-1)P_iW_i$, $b=2(\eta\_w-1)\gamma W_s$ and $c=(\eta\_w-1)\gamma^2 W_s$.

Perhaps counter-intuitively, the equation above does not predict that $\Rnum$ is monotone decreasing with respect to $\omega$. In other words; our model does not predict that returning test results more rapidly \textit{always} lower $\Rnum$. In order to gain insight into this intriguing behavior, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly as follows.
\begin{align}\label{eq:defQ}
    Q =& \frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{w}-1}P_{i}\right).
\end{align}
With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
\begin{align}
    \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
    \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}.
\end{align}

@@

Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favorable). By the previous analysis, this is equivalent to solving for $\omega > \omega_2$. So
\begin{align}
    &\omega > \omega_2 \nonumber \\
    &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
    &\sqrt{1-Q} > \frac{\gamma}{\omega}+1 \\
    &1-Q > (\frac{\gamma}{\omega}+1)^2.
\end{align}
Substituting in $Q$ from \eqref{eq:defQ} we have
\begin{align}
    &1-\frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2 \\
    &-\frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1 \\    
    &\frac{W_i}{W_s}\left(\frac{1-n_{t}}{1-n_{\omega}}P_{i}-1\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1\label{eq:necsuf}.
\end{align}
Since all steps in deriving \eqref{eq:necsuf} are reversible, \eqref{eq:necsuf} gives a necessary and sufficient condition for $\omega > \omega_2$, which characterizes when returning tests more rapidly would cause a decrease in $\Rnum$.

@@

Consider a test that allows us to test at rate $\rho_1$ and has sensitivity $P_{i,1}$, and another test that allows us to test at $\rho_2$ and has sensitivity $P_{i,2}$. Suppose that $\rho_1 > \rho_2$. Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta\_w-1)(\gamma W_s+\omega W_i) + (\eta\_c -1)P_iW_i \omega^2 \Big).$$

Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, P_{i,2}) < \Rnum(\rho_1, P_{i,1})$$ into

\begin{align}\label{eq:rho1vsrho2}
    &\rho_1\left(\gamma(\eta\_w-1)(\gamma W_s + \omega W_i) + (\eta\_c-1)P_{i, 1}W_i\omega^2\right) - \rho_2\left(\gamma(\eta\_w-1)(\gamma W_s + \omega W_i) + (\eta\_c-1)P_{i, 2}W_i\omega^2\right) > 0 \nonumber \\
    &\vdots \nonumber \\
    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta\_w}{1-\eta\_c}\cdot \frac{\gamma(\gamma W_s + \omega W_i)}{\omega^2 W_i}
\end{align}

>>>>

\begin{figure}[!h] 
\begin{center} 
\includegraphics[scale=2]{./pix/sir_comp.pdf}
\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}. Here, the disease-based status of a compartment $X$, where $X \in \{S,I,R\}$, is combined with the testing-based status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively. Also,  $\Lambda$ is the force of infection with definition in Eq.~\eqref{Lambda}, $\gamma$ is the recovery rate, $\omega$ is the rate of test return, $F_X$ and $p_X$ represent the \percap testing rate and the probability of positive tests, respectively, for compartment $X$. For further description of the parameters see \cref{tab:params}.
\label{fig:flowchart}}
\end{center} 
\end{figure}

Table~\ref{tab:params} defines the model parameters, which are generally straightforward
\percap flows between compartments, or modifiers to these flow rates. The \percap testing intensity is defined as the number of daily tests taken in a population of size $N_0$. 
The novel component of the model comes in through the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively. For example, $w_I/w_S=2$ means that infected individuals are tested at twice the \percap rate of susceptible individuals.

In order to link to more applied models, i.e., including both epidemiological and testing processes, we constructed this model so we could specify the total \percap testing rate. We do this by defining the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculating a scaling parameter for testing as:
\begin{equation}
\label{sigma}
\sigma = \frac{\rho N_0}{W}.
\end{equation}
Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma W_X$, where $X \in \{S,I,R\}$. 
For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.  Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can (if large enough) exhaust the pool of people available for testing, leading to a singularity when no one is left untested.  Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (we discuss an adjustment to the model that solves this problem in the appendix).

@@

\begin{table}[htp]
\centering
\begin{tabular}{|c|L{2in}cc|} \hline
  Symbol & Description & Unit & Value \\ \hline
  $N_0$     & Total population size & people & $10^6$ \\ \hline
  $\omega$  & Rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
  $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
  $\eta\_w$  & Relative probability of transmission for ``waiting'' individuals & - & - \\ \hline
  $\eta\_c$  & Relative probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
  $\beta$ & Transmission rate & 1/day & 0.39 \\ \hline
  $\Lambda$ & Force of infection & 1/day & - \\ \hline
  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
  $w_S, w_I, w_R$ & Relative testing weights & - &
  \begin{minipage}[t]{0.21\columnwidth}%
    Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
  \end{minipage} \\
  \hline
  \end{tabular}
\caption{\label{tab:params} Parameters of the model \eqref{model}.}
\end{table}

@@

The analytical calculation of the next generation matrix and the derivation of $\Rnum$ was carried out in Maple \citep{maple14} by using simple linear Algebra package. 
  We used R \citep{r} for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ \eqref{R0}. The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing. The parameter values or ranges are presented in Table~\ref{tab:params}. This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $w_S=w_I=w_R$, and (ii) non-random testing, when testing weight are not equal. For simulation purposes we chose $w_S=w_I=w_R=1$ for random testing, and $w_S=0.3$ and $w_I=w_R=1$ for non-random testing strategy. Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.

@@

\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_random.pdf}
\caption{}\label{p.a}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_TTI.pdf}
\caption{}\label{p.b}
\end{subfigure}
\caption{
{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at differnt levels of testing and isolation.}
We numerically evaluate $\Rnum$ \eqref{R0} using the following parameters (listed in Table \ref{tab:params}):
$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\eta\_w \in [0,1]$, $\eta\_c \in [0,1]$, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$. Contours of $\Rnum$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$. The black solid line in each panel represents the critical contour of $\Rnum=1$. \bmb{we still have some collisions in the y-axis tick labels}
% \ali{plotr + expand_limits(y = 0), change 1/\omega on x-axis}
}
\label{pan}
\end{figure}

@@

We incorporated the compartment-specific relative testing weights, $w_S$, $w_I$ and $w_R$, to model random testing and \emph{targeted} testing strategies. Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $w_I/w_S=3$ and $w_R=w_I$. Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario. 
This part needs to be developed further in future work. 
Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment. 
This can be implied when we would like to quantify and compare the effect of different levels of test focus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management. For example, when people are tested for "screening", the individuals with potential higher mobility, eg. people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with. With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing processes.

@@

\ali{ not sure if we want these 2 following parageraphs, or shrink it into afew short sentences?}
In addition to the favorable-delay-reporting observation of our model \eqref{model}, the model enables us to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$.
To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf}. To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta\_c \ll \eta\_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $w_I \gg w_S$).

@@

% %%%%%%%
\subsection{Taylor Approximation of $\Rnum$ at $\rho=0$ }
The basic reproduction number, $\Rnum$, close to $\rho=0$ can be approximated linearly in $\rho$ by using Taylor approximation. it follows
\begin{equation}
\label{eq:R0appr}
\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big) + \mathcal{O}(\rho^2).
\end{equation}

Also, to analyze the influence of $\omega$ on $\Rnum$, the approximation \eqref{eq:R0appr} was used. Then it is straight forward to have
\begin{equation}
\label{Rom}
\partial{\Rnum}/\partial{\omega}=  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
\end{equation}
where $a=(\eta\_w-1)w_I-(\eta\_c-1)p_I w_I = ((s-p_I)\eta\_c + (p_I-1)) w_I$, $b=2(\eta\_w-1)\gamma w_S$ and $c=(\eta\_w-1)\gamma^2 w_S$.
Given that $0 \leq \eta\_c\leq \eta\_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.

@@

% %%%%%%%
\subsection{rate of returning tests} 
\ali{needs editig}
The linearization of $\Rnum$ around $\rho=0$ is
\begin{equation}\label{linearization}
\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big). 
\end{equation}
So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
where $a=(\eta\_w-1)w_I-(\eta\_c-1)p_I w_I$, $b=2(\eta\_w-1)\gamma w_S$ and $c=(\eta\_w-1)\gamma^2 w_S$.

Perhaps counter-intuitively, the equation above does not predict that $\Rnum$ is monotone decreasing with respect to $\omega$. In other words; our model does not predict that returning test results more rapidly \textit{always} lower $\Rnum$. In order to gain insight into this intriguing behavior, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly as follows.
\begin{align}\label{eq:defQ}
    Q =& \frac{w_I}{w_S}\left(1-\frac{n_t-1}{n_w-1}p_I \right).
\end{align}
With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
\begin{align}
    \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
    \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}.
\end{align}

@@

Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favorable). By the previous analysis, this is equivalent to solving for $\omega > \omega_2$. So
\begin{align}
    &\omega > \omega_2 \nonumber \\
    &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
    &\sqrt{1-Q} > \frac{\gamma}{\omega}+1 \\
    &1-Q > (\frac{\gamma}{\omega}+1)^2.
\end{align}
Substituting in $Q$ from \eqref{eq:defQ} we have
\begin{align}
    &1-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2 \\
    &-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1 \\    
    &\frac{w_I}{w_S}\left(\frac{1-n_{t}}{1-n_{\omega}}P_{i}-1\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1\label{eq:necsuf}.
\end{align}
Since all steps in deriving \eqref{eq:necsuf} are reversible, \eqref{eq:necsuf} gives a necessary and sufficient condition for $\omega > \omega_2$, which characterizes when returning tests more rapidly would cause a decrease in $\Rnum$.

@@

Consider a test that allows us to test at rate $\rho_1$ and has sensitivity $P_{i,1}$, and another test that allows us to test at $\rho_2$ and has sensitivity $P_{i,2}$. Suppose that $\rho_1 > \rho_2$. Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big).$$

Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, p_{I,2}) < \Rnum(\rho_1, p_{I,1})$$ into

\begin{align}\label{eq:rho1vsrho2}
    &\rho_1\left(\gamma(\eta\_w-1)(\gamma w_S + \omega w_I) + (\eta\_c-1)p_{I,1}w_I\omega^2\right) - \rho_2\left(\gamma(\eta\_w-1)(\gamma w_S + \omega w_I) + (\eta\_c-1)p_{I, 2}w_I\omega^2\right) > 0 \nonumber \\
    &\vdots \nonumber \\
    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta\_w}{1-\eta\_c}\cdot \frac{\gamma(\gamma w_S + \omega w_I)}{\omega^2 w_I}
\end{align}
